%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% acknowledge.tex: Acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
It's a great honor to pursue my Ph.D. at the University of Minnesota. I'd like to express my deepest gratitude to all of those who have supported me throughout this amazing journey.

I am deeply grateful to my advisor, Professor Nicola Elia, for his invaluable guidance and unwavering support throughout these years. He has profoundly shaped my research taste and cultivated my habit of rigorous thinking. His insightful suggestions and constant encouragement have been essential to the completion of this work. I truly cherish the time I have spent learning under his mentorship.

I would like to thank the U.S. National Science Foundation for their financial support of my research. I am also grateful to the Department of Mechanical Engineering at the University of Minnesota for providing teaching assistant opportunities that supported my living expenses. Additionally, I would like to express my sincere appreciation to my committee members---Dr.~Rajesh Rajamani, Dr.~Murti Salapaka, and Dr.~Andrew Lamperski---for their valuable guidance and insightful suggestions throughout my research.

In addition, I consider it a great privilege and honor to have worked with so many outstanding colleagues at the University of Minnesota. I would like to thank Zaifu Zhan, Sean Yin, Dr.~Cailong Hua, Dr.~Bingqing Song and all my colleagues. It has always been a pleasure to engage in discussions with them. Their constructive feedback and ongoing encouragement have been instrumental in refining my ideas. I am especially grateful to Qingyuan Jiang, whose professional and insightful contributions to robotics applications have been offered generously and without reservation.

Last but not least, I am profoundly grateful to my family and friends. 
Their constant support has been a pillar of strength for me. 
Their belief in me has also been a source of motivation, encouraging me to face challenges and overcome difficulties. I could not have come this far without them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
